Neural Connectome of the Ctenophore Statocyst

### 2025-07-10 manuscript/Manuscript.qmd
add video7 and ed Method

To ensure consistency in gravitational input, we included only larvae whose aboral–oral axis was oriented nearly perpendicular to the gravitational vector (tilted by no more than 20°). This minimized the potential influence of larval tilt on left–right differences in balancer cilia activity, as all balancers received comparable stimulation from the statolith.

@@

Balancer cilia movement was imaged using a differential interference contrast (DIC) microscope (Zeiss Axio Imager.M2) equipped with a high-sensitivity monochrome CMOS camera optimized for the near-infrared range (UI-3360CP-NIR-GL Rev.2, iDS), and a 40x glycerine-immersion objective lens (LD LCI Plan-Apochromat 40x/1.2 Imm Corr DIC M27). To ensure consistent statolith positioning during imaging, the microscope was tilted by 90° to achieve vertical stage alignment. Illumination was provided by a custom-made NIR LED strobe system (850 nm) synchronized with the camera’s exposure signals. Image acquisition was controlled using BohNavi software. Recordings were acquired at 640 × 480 pixels, 100 frames per second (fps), for a duration of 2 minutes, using 0.05 ms pulse illumination from the LED.

@@

Traces of left and right cilia were extracted, cleaned to remove missing or non-finite values, and indexed by frame. Data were saved as RDS objects for downstream analysis. Mean values were plotted as bar graphs with the left and right cilia shown in positive and negative directions, respectively.

@@

>>>>

To ensure consistency in gravitational input, we selected only larvae whose aboral–oral axis was oriented nearly perpendicular to the gravitational vector (within 20° tilt). This minimized the potential influence of larval tilt on left–right differences in balancer cilia activity, as all balancers received comparable stimulation from the statolith.

@@

Because tilt angle strongly affects the beat frequency of balancer cilia via asymmetric mechanical loading by the statolith [@Tamm1980], our quantitative analysis was restricted to recordings that met the above orientation criterion (tilt < 20°). In this orientation, gravitational input from the statolith is approximately symmetrical across the left and right balancers, allowing meaningful comparisons of their activity. To illustrate how excessive tilt induces asymmetric stimulation, Video 7 presents two examples of larvae outside the permitted range: one tilted by ~135° in the sagittal plane and the other by ~60° in the tentacular plane. In both cases, strong asymmetries in balancer ciliary beating are evident, consistent with previous observations of gravity-induced geotactic responses [@Tamm1980]. To illustrate how excessive tilt induces asymmetric stimulation, these examples demonstrate why we restricted our analysis to near-vertical orientations, where gravitational input is approximately balanced across the balancers.

Balancer cilia movement was imaged using a differential interference contrast (DIC) microscope (Zeiss Axio Imager.M2) equipped with a 40x glycerine-immersion objective (LD LCI Plan-Apochromat 40x/1.2 Imm Corr DIC M27). Images were acquired with a high-sensitivity monochrome CMOS camera (UI-3360CP-NIR-GL Rev.2, iDS), optimized for near-infrared detection. Illumination was provided by a custom-built 850 nm LED strobe system, synchronized with camera exposure using BohNavi software. Recordings were acquired at 640 × 480 pixels, 100 frames per second (fps), for a duration of 2 minutes, using 0.05 ms pulse illumination from the LED.

@@

Traces of left and right cilia were extracted, cleaned to remove missing or non-finite values, and indexed by frame. Data were saved as RDS objects for downstream analysis. Mean values were plotted as bar graphs with left and right cilia plotted as positive and negative bars, respectively, to visually distinguish their activity.

@@

**Video 7.** High-speed imaging of balancer ciliary beating in the aboral organ tilted relative to gravity, illustrating gravity-induced asymmetry. The example on the left shows an aboral organ tilted by approximately 135° in the sagittal plane, while the example on the right shows one tilted by approximately 60° in the tentacular plane. For each case, the left and right balancer ciliary clusters are displayed separately. Ciliary beating is shown alongside kymographs extracted from two representative clusters. Recordings were acquired at 100 frames per second and are played back at 30 frames per second (0.3×). A white vertical line overlaid on each kymograph indicates the current frame shown in the video. These examples demonstrate asymmetrical balancer activity consistent with geotactic responses reported by Tamm (1980) [@Tamm1980].
